---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
```

```{r}
head(data)
```

## What is mean total number of steps taken per day?
### Import data manipulation library
```{r}
library(dplyr)
```

### Get average steps, total steps, and median for each day (ignoring the NA values)
```{r}
proc_data <- data %>% filter(!is.na(steps)) %>%
    mutate(date = as.Date(date, "%Y-%m-%d")) %>% 
    group_by(date) %>% summarise(total_steps = sum(steps), avg_steps = mean(steps))

    
head(proc_data)
```

### Import ggplot2 library for plots
```{r}
library(ggplot2)
```

### Histogram of steps 
```{r}
h <- ggplot(proc_data, aes(total_steps))
h + geom_histogram(fill = 'blue') + labs(title = "Steps per day Histogram") + 
    geom_vline(aes(xintercept = mean(total_steps), color = "mean"), size = 2) + 
    geom_vline(aes(xintercept = median(total_steps), color = "median"), alpha=0.5) +
    scale_color_manual(values = c("mean"="green", "median"="black")) #this map is added in order to get the legend 
    
```
### Average and median total steps per day
```{r}
print(paste("Mean total steps per day:", round(mean(proc_data$total_steps))))
print(paste("Median total steps per day:", round(median(proc_data$total_steps))))

```
## What is the average daily activity pattern?
```{r}
daily_pattern <- data %>% filter(!is.na(steps)) %>%
    mutate(date = as.Date(date, "%Y-%m-%d")) %>% 
    group_by(interval) %>% summarise(avg_steps = mean(steps))

head(daily_pattern)
```
```{r}
g <- ggplot(daily_pattern, aes(x = interval, y = avg_steps))
g + geom_line(color = 'blue', size = 1)

```
```{r}
print(daily_pattern[daily_pattern$avg_steps == max(daily_pattern$avg_steps),])
```
## Imputing missing values
### Report the number of missing values
```{r}
nas <- sum(is.na(data$steps))
print(paste("The number of NA values in the steps column is:", nas))
```

### Filling missing values using the mean value for the corresponding interval
```{r}
filled_data <- data %>% left_join(daily_pattern, by = "interval") %>%
    mutate(steps = if_else(is.na(steps), avg_steps , steps)) %>%
    select(steps, date, interval)

head(filled_data)
```
### Histogram of steps (wona = without NAs)
```{r}

data_wona <- filled_data %>%
    mutate(date = as.Date(date, "%Y-%m-%d")) %>% 
    group_by(date) %>% summarise(total_steps = sum(steps))

head(data_wona)
```

```{r}
h_wona <- ggplot(data_wona, aes(total_steps))
h_wona + geom_histogram(fill = 'brown') + labs(title = "Steps per day Histogram") + 
    geom_vline(aes(xintercept = mean(total_steps), color = "mean"), size = 2) + 
    geom_vline(aes(xintercept = median(total_steps), color = "median"), alpha=0.5) +
    scale_color_manual(values = c("mean"="green", "median"="black")) #this map is added in order to get the legend 
    
```
### Average and median total steps per day (without NAs)
```{r}
print(paste("Mean total steps per day:", round(mean(data_wona$total_steps))))
print(paste("Median total steps per day:", round(median(data_wona$total_steps))))

```
### There is apreciable change in mean nor median nor in the histogram. The count of steps is greater because before we were ignoring NA values and now there is a mean count filling those missing values. 

## Are there differences in activity patterns between weekdays and weekends?
```{r}

data_weekdays <- filled_data %>%
    mutate(date = as.Date(date)) %>%
    mutate(weekday = if_else((weekdays(date) %in% c("Saturday", "Sunday")), "weekend", "weekday")) %>%
    group_by(interval, weekday) %>%
    summarise(avg_steps = mean(steps))

head(data_weekdays)
```

### Plot the total average by interval and weekday-weekend category.
```{r}
q <- ggplot(data_weekdays, aes(x = interval, y = avg_steps))

q + geom_line(color = "blue") + facet_grid(rows = vars(weekday))

```