---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
```

```{r}
head(data)
```

## What is mean total number of steps taken per day?
###Import data manipulation library
```{r}
library(dplyr)
```

### Get average steps, total steps, and median for each day (ignoring the NA values)
```{r}
proc_data <- data %>% filter(!is.na(steps)) %>%
    mutate(date = as.Date(date, "%Y-%m-%d")) %>% 
    group_by(date) %>% summarise(total_steps = sum(steps), avg_steps = mean(steps))

    
head(proc_data)
```

###Import ggplot2 library for plots
```{r}
library(ggplot2)
```

###Histogram of steps 
```{r}
h <- ggplot(proc_data, aes(total_steps))
h + geom_histogram(fill = 'blue') + labs(title = "Steps per day Histogram") + 
    geom_vline(aes(xintercept = mean(total_steps), color = "mean"), size = 2) + 
    geom_vline(aes(xintercept = median(total_steps), color = "median"), alpha=0.5) +
    scale_color_manual(values = c("mean"="green", "median"="black"))
    
```
### Average and median total steps per day
```{r}
print(paste("Mean total steps per day:", round(mean(proc_data$total_steps))))
print(paste("Median total steps per day:", round(median(proc_data$total_steps))))

```
## What is the average daily activity pattern?
```{r}
daily_pattern <- data %>% filter(!is.na(steps)) %>%
    mutate(date = as.Date(date, "%Y-%m-%d")) %>% 
    group_by(interval) %>% summarise(avg_steps = mean(steps))

head(daily_pattern)
```
```{r}
g <- ggplot(daily_pattern, aes(x = interval, y = avg_steps))
g + geom_line(color = 'blue', size = 1)

```
```{r}
print(daily_pattern[daily_pattern$avg_steps == max(daily_pattern$avg_steps),])
```
## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
